---
title: "The prime of their life - the halflife of acting talent"
output: 
html_document:
code_folding: hide
---

> insert an image

Much analysis has been completed regarding the performing peak of athletes and sportspeople. Statistically, swimmers peak at around 24, marathon runners between 30-31, gymnasts at a mere 18, NBA players generally peak between 26-29, professional hockey players between 27-28, and chess players at 30-35. 

Does the idea of a performing prime extend to artists, specifically actors and directors? There are obvious reasons to be skeptical. Importantly there's no major physical component to acting that would decay as the body does, and unlike chess you wouldn't think there's the need for an exreme state of concentration and cognition which might be unreachable at older ages. 

There are significant confounding factors that might prevent us from answering this question difinitively (which I'll discuss later) but let's dig into the data.

# Housekeeping

Let's set ourselves up with some packages, themes and utility functions.

```{r}
# Packages
library(readr)
library(dplyr)
library(tidyr)
library(gender)
library(stringr)
library(lubridate)
library(highcharter)

# Rmarkdown Options
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

# For reproducibility
set.seed(123)

# Convert an R vector to string JS vector
convert_js_vec = function(vec) {
  js_string = vec %>% round(4) %>% paste(collapse = ",")
  return(JS( paste0("[", js_string, "]") ))
}

# Highcharts theme
source("https://raw.githubusercontent.com/TMBish/lab_posts/master/highcharts_theme.R")
tmbish = high_tmbish()
```

# Data

To curate a data set for anlaysis I started with imdb user **hagennelson's** list of [top 1000 actors](http://www.imdb.com/list/ls058011111/) and imdb user **earthanime's** list of [top 250 directors](http://www.imdb.com/list/ls062419846/). Both lists span across generations, from the ancient (Ethel Barrymore and Alfred Hitchcock) to the modern (Chloë Grace Moretz and Christopher Nolan).

I then developed code (available on GitHub) to scrape data from the rotten tomatoes pages of the above individuals' films. The final dataset includes **12,258** films and **1,233** indviduals. Although we'll only use a few of these variables, I brought in a few extras for some future investigations:

Variable | Description
------------ | ---------------------
url | URL extension on rotten tomatoes website
year | Film production year
title | Film title
reviews | Count of formal critic reviews of film
tomatometer | Percentage of favourable reviews
av_rating | Average rating of film out of 10
audience | The non-critic plebian equivolent of the tomatometer
box_office | The box-office revenue
director | The film's director
actor_1:actor_6 | The 6 top actors according to rotten tomatoes internal ranking (seemingly ordered by screen time)

## Film Data

First we'll load and take a peek at the film dataset:

```{r}
# Load in films data from Github Repo
films = read_csv("https://raw.githubusercontent.com/TMBish/lab_posts/master/01_actor_careers/00_full_film_data.csv") %>% filter(!is.na(title))

films %>% head()

```

## Individual Data

Now we'll load the individual data. 

Unfortunately, because the Rotten Tomatoes development tea are a bunch of politically correct SJWs, gender isn't provided on their film bio pages so we'll have to be creative. The `gender` R package allows you to infer a person's gender from their first name and DOB. The result is a probabilistic estimate of that individuals most likely gender given their first name and DOB.

```{r}

library(gender)
library(stringr)
library(lubridate)

film_folk = read_csv("https://raw.githubusercontent.com/TMBish/lab_posts/master/01_actor_careers/00_actors_directors.csv")

# Clean some data quality issues on the DOB
film_folk[film_folk$dob < as.Date("1880-01-01"), "dob"] = as.Date("1880-01-01")

# Grep out first name and dob year for the gender function
film_folk = film_folk %>%
              mutate(
                first_name = str_extract(name, "(?i)[a-z]+(?=\\s)"),
                dob_year = year(dob)
              )

# Distinct name, dob combos and predict gender
genders = film_folk %>%
            distinct(first_name, dob_year) %>%
            gender_df(name_col = "first_name", year_col = "dob_year") %>%
            select("first_name" = name, gender, "dob_year" = year_min)

```

We should sense check the gender mapping to make sure the package is making logical inferences. Let's take a random 10 person sample:

```{r}
genders %>%
  sample_n(10) %>%
  knitr::kable()
```

So generally this is looking pretty, pretty good. Things get dicey with unisex names like Sandy; however, the package must know that most Sandys born in 1937 were female so we'll have to trust it here. Finally we'll need to join this information back to the individual data set.

```{r}
# Join back the gender back to the main data
film_folk = 
  film_folk %>%
  left_join(genders)

```

# What's the deal with female directors?

To add a layer to this analysis I wanted group people by gender and film production role (actor vs director). So we think we'd be working with 4 groups, however, I was suprised to see that of our 250 top directors only XXX were female. A consequence of history but appaling nonetheless.

This got me thinking about the rest of the dataset. Given our **12,258** film dataset isn't limited to director credits in our top 250 (films in the set are contributed by both actor and director credits) we can get a bit of a better idea of the broader gender split of directors. I've captured a lot of the key numbers in the following chart.

```{r}
# Establish a persistent dataset of interest
data = films %>%
        select(title, year, tomatometer, director:actor_6) %>%
        gather("role", "name", director:actor_6) %>%
        inner_join(film_folk) %>%
        select(-first_name) %>%
        mutate(
          role_type = ifelse(role == "director", "Director", "Actor"),
          age_at_production = year - dob_year,
          category = case_when(
            role_type == "Director" ~ "director",
            gender == "male" ~ "male_actor",
            gender == "female" ~ "female_actor",
            TRUE ~ "other"
          )
        )


# Directors

# Data for known individuals
known_summary = 
  data %>%
  group_by(role_type, gender) %>%
  summarise(
    films = n(),
    distinct_people = n_distinct(name)
  )

# Unknown directors
unknown_set = films %>%
          select("name" = director) %>%
          anti_join(film_folk) %>%
          mutate(
            first_name = str_extract(name, "(?i)[a-z]+(?=\\s)"),
            min_year = 1900, max_year = 1980
          ) %>%
          filter(!is.na(first_name))

unknown_genders = 
  unknown_set %>% 
  gender_df(name_col = "first_name", year_col = c("min_year", "max_year")) %>%
  select("first_name" = name, gender)

unknown_summary = 
  unknown_set %>%
  inner_join(unknown_genders) %>%
  mutate(role_type = "Director (Non Top250)") %>%
  group_by(role_type, gender) %>%
  summarise(
    films = n(),
    distinct_people = n_distinct(name)
  )
  
chart_data = 
  known_summary %>%
  union_all(unknown_summary) %>%
  filter(!is.na(gender))


map(c("films", "distinct_people"), function(x) {

    data = 
      chart_data %>%
      select_("role_type", "gender", x) %>%
      spread_("gender", x, fill = 0)
      
    base = highchart() %>%
            hc_chart(type = "column") %>%
            hc_add_theme(tmbish) %>%
            hc_xAxis(categories = data$role_type) %>%
            hc_plotOptions(
              column = list(
                stacking = 'percent',
                dataLabels = list(
                  enabled = TRUE
                  #,formatter = JS("function(){ return(point.percentage)}")
                )
              )
            ) %>%
            hc_title(text = ifelse(x=="films", "By Volume of Films", "By Distinct Individuals"))

    for (i in c("female", "male")) {
      base = base %>% hc_add_series(name = i, data = data[,i] %>% unlist() %>% as.integer())
    }
    
    if (x == "films") { base  = base %>% hc_legend(enabled=FALSE)}
    
    return(base)
  
}) %>%
hw_grid(ncol = 2) %>% 
browsable()

```

# How old are we when we produce films? 



We've finally sorted out our source data. However, we still need to add a few variables to be able to answer this question. We need we'll first need to calculate a person's age at the time of their appearance in a film. We'll then 

```{r}

# Chain a bunch of data wrangling operations
volume_data =
  films %>% 
  select(title, year, tomatometer, director, actor_1:actor_6) %>%
  gather("role", "name", director, actor_1:actor_6) %>%
  mutate(role = ifelse(role == "director", "Director", "Actor")) %>%
  inner_join(film_folk) %>%
  mutate(
    age_at_production = year - dob_year
  ) %>%
  filter(
    between(age_at_production, 10,100),
    !is.na(gender)
  ) %>%
  select(name, gender, dob_year, age_at_production, year, role, tomatometer) %>%
  mutate(
    category = case_when(
      role == "Director" ~ "director",
      gender == "male" ~ "male_actor",
      TRUE ~ "female_actor"
    )
  ) %>% 
  group_by(category, age_at_production) %>%
  summarise(
    films = n()
  ) %>%
  group_by(category) %>%
  mutate(
    percentage = films / sum(films)
  )

```

<div id = "age_series_chart">
```{r}

# Shape data for presentation
chart_data = 
  volume_data %>%
  select(-films) %>%
  spread(category, percentage) %>%
  mutate(
    director = ifelse(!is.na(director), director, 0),
    female_actor = ifelse(!is.na(female_actor), female_actor, 0),
    male_actor = ifelse(!is.na(male_actor), male_actor, 0),
    director_cdf = cumsum(director),
    female_actor_cdf = cumsum(female_actor),
    male_actor_cdf = cumsum(male_actor)
  )


# For the javascript client side interactivity on the age series chart
female_cdf = chart_data$female_actor_cdf %>% convert_js_vec()
female_pdf = chart_data$female_actor %>% convert_js_vec()
male_cdf = chart_data$male_actor_cdf %>% convert_js_vec()
male_pdf = chart_data$male_actor %>% convert_js_vec()
dir_cdf = chart_data$director_cdf %>% convert_js_vec()
dir_pdf = chart_data$director %>% convert_js_vec()

# Create chart
age_dist_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(
    animation = list(duration = 2000)
  ) %>%
  hc_xAxis(
    categories = chart_data$age_at_production,
    title = list(text = "Age at Production")
  ) %>%
  hc_yAxis(
    title = list(text = "Probablity Density"),
    labels = list(formatter = JS("function(){return(Math.round(this.value * 1000) / 10 + '%')}"))
  ) %>%
  hc_title(text = "Female actors") %>%
  hc_subtitle(text = "Comparing male actors, (male) directors and female actors") %>%
  hc_add_series(
    name = "Director", 
    type = "spline",
    data = chart_data$director,
    marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Male Actor",
    type = "spline",
    data = chart_data$male_actor,
    marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Female Actor",
    type = "area",
    data = chart_data$female_actor,
    zIndex = -10,
    marker = list(enabled = FALSE),
    fillOpacity = 0.5
  )

age_dist_chart
```
</div>

<button class="button secondary outline" id = "update_age_series"> Show Cumulative Distribution </button>

# Q2. Acting and Directing as a skill

> Hypothesis 2: acting and directing skills aren't tied to a person's physical attributes so shouldn't experience deterioration in age. In fact you would expect these artists to improve their skills with age.

## Measuring Performance

To investigate our hypothesis we must make a couple of assumptions. Since we don't have an individual actor / director performance (like we have in sports) we'll have to assume an actor or director can have a tangible affect on the quality of the movie they appear in; there are obviously many other factors but this seems like a relatively safe assumption.

Further - for actors - we'll have to limit the dataset to actors that played an important role in the film. To implement an `actor performance == film quality` approach the actors need to be prominent enough so as to be able to affect the overall quality of the film. So while Michael Cera's cameo in **This is the End** was high quality it can't be said that he influenced the rotten tomatoes rating of the film on his own.

<iframe width="560" height="315" src="https://www.youtube.com/embed/qIKPJlKHKxg" frameborder="0" allowfullscreen></iframe>

After looking through a few rotten tomatoes pages it seems limiting to the top 3 actors in a given film is a reasonable cut-off for influncial performances.

> INSERT EXAMPLES FROM A FEW ROTTEN TOMATOES PAGES

## Methodology

For every person in the dataset we follow the following recipe:

1. Produce a dataset for each actor with films that:
+ We have a tomatometer score for
+ And the actor is in the top 3 of rotten tomatoes credits
2. Calculate the `mean` tomatometer for all of this person's films;
3. Calculate the age the person was at the release of each film;
4. For every film calculate what `percentage %` this tomato rating is above or below this person's `mean` (called a tomato index)

Producing a dataset like this:

```{r echo=FALSE}
films %>% 
  select(title, year, tomatometer, director, actor_1:actor_3) %>%
  gather("role", "name", director, actor_1:actor_3) %>%
  inner_join(film_folk) %>%
  filter(
    name == "Denzel Washington"
  ) %>%
  mutate(
    age_at_production = year - dob_year
  ) %>%
  filter(between(age_at_production, 15,85)) %>%
  select(title, name, year, gender, age_at_production, tomatometer) %>%
  group_by(name) %>% 
  mutate(
    average_tomato  = round(mean(tomatometer))
  ) %>% 
  ungroup() %>%
  mutate(
    tomato_index = tomatometer / average_tomato
  ) %>%
  arrange(age_at_production) %>%
  select(title, name, year, age_at_production, tomatometer, average_tomato, tomato_index) %>%
  mutate(index = scales::percent(tomato_index)) %>%
  sample_n(10) %>%
  knitr::kable()



```

## Results

```{r}
# 
rating_set =
  films %>% 
  select(title, year, tomatometer, director, actor_1:actor_3) %>%
  gather("role", "name", director, actor_1:actor_3) %>%
  inner_join(film_folk) %>%
  filter(!is.na(gender)) %>%
  mutate(
    age_at_production = year - dob_year,
    category = case_when(
      role == "director" ~ "director",
      gender == "male" ~ "male_actor",
      TRUE ~ "female_actor"
    )
  ) %>%
  filter(between(age_at_production, 15,85)) %>%
  select(name, gender, age_at_production, category, tomatometer) %>%
  group_by(name) %>% 
  mutate(
    average_tomato  = median(tomatometer),
    films = n()
  ) %>% ungroup() %>%
  filter(films >= 5) 

rating_volumes = 
  rating_set %>%
  group_by(age_at_production) %>%
  summarise(films = -1*n())
  
  
rating_scores = 
  rating_set %>%
  mutate(
    tomato_index = tomatometer / average_tomato) %>%
  group_by(category, age_at_production) %>%
  summarise(rating = median(tomato_index)) %>%
  spread(category, rating)

# LOESS Trend Line
loess_fit = loess(value ~ age_at_production , 
                  data = rating_scores %>% gather("role", "value", -age_at_production)
                  # %>% group_by(age_at_production) %>% summarise(value = mean(value, na.rm = TRUE))
)

predictions = loess_fit %>% predict(seq(15,85))


volume_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(height = "300") %>%
  hc_xAxis(categories = rating_volumes$age_at_production, opposite = TRUE) %>%
  hc_yAxis(
    title = list(text = "Total Films in Dataset"),
    labels = list(formatter = JS("function(){return(-1 * this.value)}"))    
    ) %>%
  hc_title(text = "") %>%
  hc_add_series(
    name = "Volumes",
    type = "area",
    data = rating_volumes$films,
    marker = list(enabled = FALSE)
  ) %>% 
  hc_legend(enabled = FALSE)

rating_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(
    animation = list(duration = 2000)
  ) %>%
  hc_xAxis(
    categories = rating_scores$age_at_production,
    title = list(text = "Age at Production")
  ) %>%
  hc_yAxis(
    title = list(text = "Probablity Density"),
    min = 0.6
  ) %>%
  hc_title(text = "Female actors") %>%
  hc_subtitle(text = "Comparing male actors, (male) directors and female actors") %>%
  hc_add_series(
    name = "Director",
    type = "scatter",
    data = rating_scores$director,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Male Actor",
    type = "scatter",
    data = rating_scores$male_actor,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Female Actor",
    type = "scatter",
    data = rating_scores$female_actor,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Trend",
    type = "spline",
    data = predictions,
    color = "#000000",
    marker = list(enabled = FALSE)
  )

```

```{r echo = FALSE}
rating_chart
```

```{r echo = FALSE}
volume_chart
```


<script>

$(document).ready(function() {

var show_cdf = false;

$("#update_age_series").click(function(){

var chart=$("#age_series_chart").find(".highchart").highcharts();

if (show_cdf) {

chart.yAxis[0].setExtremes(0,null);
chart.yAxis[0].setTitle({text:"Probability Density", redraw:false});
dir_pdf = `r dir_pdf`;
female_pdf = `r female_pdf`;
male_pdf = `r male_pdf`;
chart.series[0].setData(dir_pdf);
chart.series[1].setData(male_pdf);
chart.series[2].setData(female_pdf);
$("#update_age_series").html("Show Cumulative Distribution");

} else {

chart.yAxis[0].setExtremes(0,1);
chart.yAxis[0].setTitle({text:"Cumulative Probability", redraw:false});
dir_cdf = `r dir_cdf`;
female_cdf = `r female_cdf`;
male_cdf = `r male_cdf`;
chart.series[0].setData(dir_cdf);
chart.series[1].setData(male_cdf);
chart.series[2].setData(female_cdf);

$("#update_age_series").html("Show Density");

};

show_cdf = !show_cdf

}); 
});

</script>
